---
title: "analysis"
author: "Zheyan Liu"
date: "3/18/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .7,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Data Processing

## Drop 2% head and tail data

Read data and only keep fuel = 'electric'

```{r pressure, echo=FALSE}
df = read_csv('data/vehicles.csv', show_col_types = FALSE)
df = 
  df %>% 
  filter(fuel == 'electric') 
```

## Missing variables

### Missing rate

```{r}
as.data.frame(sapply(X = df, FUN = function(x) sum(is.na(x)))/nrow(df))  %>% knitr::kable(col.names = "Missing rate")
```

### MAR or MNAR

Impute as ‘NAN’ for analysis

```{r}
df1 = 
  df %>% 
    select(price, year, manufacturer, model, condition, odometer, title_status, transmission,drive, type, paint_color, state) %>% 
    mutate(
      manufacturer = ifelse(is.na(manufacturer), 'NAN', manufacturer),
      model = ifelse(is.na(manufacturer), 'NAN', manufacturer),
      condition = ifelse(is.na(condition), 'NAN', condition),
      odometer = ifelse(is.na(odometer), mean(odometer, na.rm=TRUE), odometer),
      title_status = ifelse(is.na(title_status), 'NAN', title_status),
      transmission = ifelse(is.na(transmission), 'NAN', transmission),
      drive = ifelse(is.na(drive), 'NAN', drive),
      type = ifelse(is.na(type), 'NAN', type),
      paint_color = ifelse(is.na(paint_color), 'NAN', paint_color)
    )

```


Missing at random (MAR) or missing not at random (MNAR)? Or in other words, are they dependent on price? Only on variable with considerably large missing rate.

#### Using box plot with price

Manufacturer is MNAR

```{r}
df1 %>% 
  group_by(manufacturer) %>% 
  mutate(n = n()) %>% 
  filter(n >= 10) %>% 
  ggplot(aes(x=manufacturer, y=price, fill=manufacturer)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
model is MNAR

```{r}
df1 %>% 
  group_by(model) %>% 
  mutate(n = n()) %>% 
  filter(n >= 30) %>% 
  ggplot(aes(x=model, y=price, fill=model)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Condition is MAR

```{r}
df1 %>% 
  group_by(condition) %>% 
  mutate(n = n()) %>% 
  filter(n >= 100) %>% 
  ggplot(aes(x=condition, y=price, fill=condition)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

title_status is MNAR

```{r}
df1 %>% 
  group_by(title_status) %>% 
  mutate(n = n()) %>% 
  filter(n >= 5) %>% 
  ggplot(aes(x=title_status, y=price, fill=title_status)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

transmission is MAR

```{r}
df1 %>% 
  group_by(transmission) %>% 
  mutate(n = n()) %>% 
  filter(n >= 5) %>% 
  ggplot(aes(x=transmission, y=price, fill=transmission)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

drive is MAR

```{r}
df1 %>% 
  group_by(drive) %>% 
  mutate(n = n()) %>% 
  filter(n >= 100) %>% 
  ggplot(aes(x=drive, y=price, fill=drive)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

type is MAR

```{r}
df1 %>% 
  group_by(type) %>% 
  mutate(n = n()) %>% 
  filter(n >= 100) %>% 
  ggplot(aes(x=type, y=price, fill=type)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

paint_color is MAR

```{r}
df1 %>% 
  group_by(paint_color) %>% 
  mutate(n = n()) %>% 
  filter(n >= 100) %>% 
  ggplot(aes(x=paint_color, y=price, fill=paint_color)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Create new attribute 'NAN' for MNAR

```{r}
df1 = 
  df %>% 
    select(price, year, manufacturer, condition, odometer, title_status, transmission,drive, type, paint_color, state, lat , long) %>% 
    filter(
      price > quantile(price,0.05),
      price < quantile(df$price,0.99)
    ) %>% 
    mutate(
      manufacturer = ifelse(is.na(manufacturer), 'NAN_cat', manufacturer),
      # odometer = ifelse(is.na(odometer), mean(odometer, na.rm=TRUE), odometer),
      title_status = ifelse(is.na(title_status), 'NAN_cat', title_status)
      # year = ifelse(is.na(year), mean(year, na.rm=TRUE), year),
    ) %>% 
    mutate(
      condition = as.factor(condition),
      condition = fct_reorder(condition, price, median),
      
      title_status = as.factor(title_status),
      title_status = fct_reorder(title_status, price, median),
      
      manufacturer = as.factor(manufacturer),
      transmission = as.factor(transmission),
      drive = as.factor(drive),
      type = as.factor(type),
      paint_color = as.factor(paint_color),
      state = as.factor(state)
           ) 
```


# Analysis

Analysis are done without imputing MAR

## Dependent variable price

Skewed to the left with two vertices

```{r}
df %>% 
  ggplot(aes(x=price)) + geom_histogram(bins=35)
```
Only tesla

```{r}
df %>% 
  filter(manufacturer=='tesla') %>% 
  ggplot(aes(x=price)) + geom_histogram(bins=35)
```

## Price with continuous variables


```{r}
library(patchwork)

p1 = 
  df1 %>% 
    ggplot(aes(x=year, y=price)) + geom_point()

p2 = 
  df1 %>% 
    ggplot(aes(x=odometer, y=price)) + geom_point()

p1 + p2
```




## Price with category variables



```{r, fig.height = 25, fig.width = 8}
p1 = 
  df1 %>% 
    group_by(paint_color) %>% 
    mutate(n = n()) %>% 
    drop_na(paint_color) %>% 
    filter(n >= 50) %>% 
    ggplot(aes(x=reorder(paint_color, price, median, na.rm = TRUE), y=price, fill=paint_color)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="paint_color")

p2 = 
  df1 %>% 
    group_by(type) %>% 
    mutate(n = n()) %>% 
    drop_na(type) %>% 
    filter(n >= 50) %>% 
    ggplot(aes(x=reorder(type, price, median, na.rm = TRUE), y=price, fill=type)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="type")

p3 = 
  df1 %>% 
    group_by(manufacturer ) %>% 
    mutate(n = n()) %>% 
    drop_na(manufacturer ) %>% 
    filter(n >= 50) %>% 
    ggplot(aes(x=reorder(manufacturer , price, median, na.rm = TRUE), y=price, fill=manufacturer)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="manufacturer ")

p4 = 
  df1 %>% 
    group_by(condition) %>% 
    mutate(n = n()) %>% 
    drop_na(condition) %>% 
    filter(n >= 50) %>% 
    ggplot(aes(x=reorder(condition , price, median, na.rm = TRUE), y=price, fill=condition)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="condition")

p5 = 
  df1 %>% 
    group_by(drive) %>% 
    mutate(n = n()) %>% 
    drop_na(drive) %>% 
    filter(n >= 50) %>% 
    ggplot(aes(x=reorder(drive  , price, median, na.rm = TRUE), y=price, fill=drive)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="drive ")

p6 = 
  df1 %>% 
    group_by(transmission) %>% 
    mutate(n = n()) %>% 
    drop_na(transmission) %>% 
    filter(n >= 30) %>% 
    ggplot(aes(x=reorder(transmission, price, median, na.rm = TRUE), y=price, fill=transmission)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="transmission")

p7 = 
  df1 %>% 
    group_by(state) %>% 
    mutate(n = n()) %>% 
    drop_na(state) %>% 
    filter(n >= 30) %>% 
    ggplot(aes(x=reorder(state, price, median, na.rm = TRUE), y=price, fill=state)) + 
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    theme(legend.position = "none") + 
    labs(y="price", x="state")

(p1 + p2 + p3)/(p4 + p5+ p6)/p7

```

### Lat and Long

```{r, warning=FALSE}
library(leaflet)
qpal <- colorQuantile("YlOrRd", df1$price, n = 4)

df1 %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircles(lng = ~long, lat = ~lat, weight = 1, stroke = FALSE, radius = ~18000, color = ~qpal(price), opacity = 0.75, fillOpacity = 0.75) %>%
  addLegend("topright", pal = qpal, values = ~price, 
            title = "Price quantile", opacity = 0.75) %>% 
  setView(-97, 40, zoom = 4)
```


# Model

## Impute with mode

Impute with mode for category variables, median for continuous variables, MinMaxScalar for continuous variable


```{r}
calc_mode <- function(x){
  
  # List the distinct / unique values
  distinct_values <- unique(x)
  
  # Count the occurrence of each distinct value
  distinct_tabulate <- tabulate(match(x, distinct_values))
  
  # Return the value with the highest occurrence
  distinct_values[which.max(distinct_tabulate)]
}

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}


df1 = 
  df1 %>% 
  mutate(
    # category impute
    condition = replace_na(condition, calc_mode(df %>% filter(!is.na(condition))  %>%  pull(condition))),
    title_status = replace_na(title_status, calc_mode(df %>% filter(!is.na(title_status))  %>%  pull(title_status))),
    transmission = replace_na(transmission, calc_mode(df %>% filter(!is.na(transmission))  %>%  pull(transmission))),
    drive = replace_na(drive, calc_mode(df %>% filter(!is.na(drive))  %>%  pull(drive))),
    type = replace_na(type, calc_mode(df %>% filter(!is.na(type))  %>%  pull(type))),
    paint_color = replace_na(paint_color, calc_mode(df %>% filter(!is.na(paint_color))  %>%  pull(paint_color))),
    
    # continuous impute
    odometer = ifelse(is.na(odometer), median(odometer, na.rm=TRUE), odometer),
    year = ifelse(is.na(year), median(year, na.rm=TRUE), year),
    lat = ifelse(is.na(lat), median(lat, na.rm=TRUE), lat),
    long = ifelse(is.na(long), median(long, na.rm=TRUE), long),
    
    # continuous MinMaxScalar
    odometer = normalize(odometer),
    year = normalize(year),
    lat = normalize(lat),
    long = normalize(long)) %>% 
    filter(
    price > quantile(df1$price, 0.02),
    price < quantile(df1$price, 0.98)
    )

head(df1) %>% knitr::kable()
```


## Regression (Not Used for model!)

Delete all na records, get adjusted R-square 0.8974

```{r}
reg = lm(price~., data = df1)
summary(reg)
```

## Lasso


Check missing rate to ensure no missing values


```{r}
as.data.frame(sapply(X = df1, FUN = function(x) sum(is.na(x)))/nrow(df1))  %>% knitr::kable(col.names = "Missing rate")
```

Fit Lasso model

```{r}
set.seed(777)
lasso.fit <- train(price~.,
                   data = df1,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(5, -2, length=300))),
                   trControl = trainControl(method = "cv"))

lasso.fit2 <- train(price~.,
                   data = df1,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(8, -1, length=10))),
                   trControl = trainControl(method = "cv"))
```


Best tune parameters

```{r}
lasso.fit$bestTune %>% knitr::kable()
```

Coefficients

```{r}
coef = coef(lasso.fit$finalModel, lasso.fit$bestTune$lambda)
coef[,1][coef[,1] != 0] %>% knitr::kable(col.names = "Non-zero Coefficients")
```

## Regression Tree

```{r}
rpartTune <- train(price~.,
                   data = df1,
                   method = "rpart2", 
                   tuneLength = 10, 
    trControl = trainControl(method = "cv"))
plot(rpartTune)
```

```{r}
library(rpart)
rpartTree <- rpart(price~., data = df1, maxdepth = 5)
rpartTree
```


## GBDT

```{r,warning=FALSE}
library(gbm)
grid<-expand.grid(.n.trees=seq(200,500,by=100),.interaction.depth=seq(1,5,by=1),.shrinkage=c(.05,.1),
                  .n.minobsinnode=10)


gbm.train<-train(price~.,data=df1,method='gbm',trControl=trainControl(method = "cv"),tuneGrid=grid)


gbm.train$results
```



# Conclusion


## Prediction

RMSE

```{r}
set.seed(777)
res = 
  resamples(list(LASSO = lasso.fit,
                 LASSO2 = lasso.fit2,
                 TREE = rpartTune,
                 GBM2 = gbm.train))
summary(res)


mean(df$price)
```



