---
title: "2021 Craigslist used electric cars price Prediction"
author: "Zheyan Liu (zl3119)"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
--- 

\newpage


```{r setup, include=FALSE}
library(tidyverse)
library(caret)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .7,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Introduction

```{r, include=FALSE}
df_raw = read_csv('data/vehicles.csv', show_col_types = FALSE)
df = 
  df_raw %>% 
  filter(fuel == 'electric') %>% 
  filter(
      # price > quantile(price,0.02),
      # price < quantile(price,0.98)
    )

```


With the soaring oil price, more and more people are considering buying an electric car to save money. We would like to build models to help predict the used electric car price so that customers can use this model to determine if the deal is reasonable. 


We used data from [Craigslist](https://newyork.craigslist.org/), which is the world's largest collection of used vehicles for sale. The [original data](https://www.kaggle.com/datasets/austinreese/craigslist-carstrucks-data) contains price of used car from Apr 2011 to May 2011, it contains `r nrow(df_raw)` observations and  `r ncol(df_raw)` variables. Since we are only interested in cars fueled by electricity, data is reduced to `r nrow(df)` observations.


## Research Questions



## Data preparation



# Exploratory analysis


# Models


# Conclusions

